Data preparation

### 2023-02-15 04. Analysis/scripts/01_data_preparation_roster.Rmd
Create 01_data_preparation_roster.Rmd
Redirected to roster because roster data may help with duplicates in daily check in

---
title: "Data preparation"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Occupancy Patterns"
  data_folder_1: "Occupancy Data Pull"
  unique_identifier_1: !r c("employee_name","date")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("arsenal")
devtools::install_github(repo = "GSA/oesrrr")
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
```

@@

@@

```{r}
files <- 
  list.files(file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1), pattern = ".csv", full.names = TRUE)
files
```

@@

```{r}
prep_base <- files %>% 
  map_dfr(~mutate(.data = read_csv(., 
    col_types = cols(Date = col_date(format = "%m_%d_%Y"))), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct()
```

@@

@@

>>>>

---
title: "Data preparation"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Occupancy Patterns"
  data_folder_1: "Rosters"
  unique_identifier_1: !r c("snapshot_date","employee_e_mail_address")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("arsenal")
# devtools::install_github(repo = "GSA/oesrrr")
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
```

@@

## Functions

```{r}
filter_unique <- function(in_data, ...)
{
  in_data %>% 
  filter(!(if_any(c(...),~is.na(.))))
}
```

@@

```{r}
files <- 
  list.files(file.path(params$google_drive_home,"03. Data Collection",params$data_folder_1), pattern = ".xlsx", full.names = TRUE)
files
```

@@

```{r}
prep_base <- files %>% 
  map_dfr(~mutate(.data = read_xlsx(path = ., 
                                    col_types = c("date", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "date", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text")), 
        file_name = str_remove(.,pattern = file.path(params$google_drive_home,"03. Data Collection"))
      )) %>% 
  clean_names() %>% 
  distinct()
```

@@

## Filter null unique ids

```{r}
prep_base <- 
  prep_base %>% 
  filter_unique(params$unique_identifier_1)
```

@@

No duplication found.

## Addressing blanks

```{r}
blank_check <- function(in_data,in_variable_name)
{
  if(prep_base %>% select(!!in_variable_name) %>% pull() %>% class() %>% head(1) == "POSIXct")
    {
    out <- FALSE
  }

  if(prep_base %>% select(!!in_variable_name) %>% pull() %>% class() %>% head(1) != "POSIXct")
   {
      check_01 <- 
  in_data %>% 
    filter(!!sym(in_variable_name) == "") %>% 
  distinct()

out <- nrow(check_01) > 0 
   }

```{r}
blanks_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~blank_check(in_data = prep_base,.))

blanks_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    blanks_by_variable = blanks_by_variable
  )

Create dataframe of only variables that have blanks.

```{r}
blanking_variables <- 
  column_meta_data %>% 
  filter(blanks_by_variable)

blanking_variables
```

No variables with blank entries.

## Addressing nulls

```{r}
null_check <- function(in_data,in_variable_name)
{
  check_01 <- 
  in_data %>% 
    filter(is.na(!!sym(in_variable_name))) %>% 
  distinct()

```{r}
nulls_by_variable <- 
  column_meta_data %>% 
  pull(variable) %>% 
  map_lgl(~null_check(in_data = prep_base,.))

nulls_by_variable
```

Add this to the `column_meta_data` dataframe.

```{r}
column_meta_data <- 
  column_meta_data %>% 
  add_column(
    nulls_by_variable = nulls_by_variable
  )

Create dataframe of only variables that have blanks.

```{r}
null_variables <- 
  column_meta_data %>% 
  filter(nulls_by_variable)

null_variables
```

```{r}
null_variables <- 
  null_variables %>% 
  mutate(group = 
           case_when(
             variable == "supervisor_display_name" ~ "supervisor_display_name",
             variable == "supervisor_e_mail_address" ~ "supervisor_display_name",
             variable == "business_line_2_letter_department_id" ~ "business_line_2_letter_department_id",
             variable == "location_city" ~ "business_line_2_letter_department_id",
             variable == "location_state" ~ "business_line_2_letter_department_id",
             variable == "position_category_code" ~ "position_category_code",
             variable == "position_category_description" ~ "position_category_code",
             TRUE ~ variable
             )
         )

null_variables
```

```{r}
null_groups <- 
  null_variables %>% 
  select(group) %>% 
  distinct() %>% 
  pull()

null_groups
```

### Nulls removal

Removing nulls produced by:

```{r}
null_variable <- 
  null_variables %>% 
  filter(group == null_groups[step_n])

null_variable
```

Reviewing blanks.

```{r}
null_01 <- 
  prep_base %>% 
  filter(if_any(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(all_of(null_variable %>% pull(variable)),everything())

null_01
```

Not too many.

For those with null `supervisor_display_name` and `supervisor_e_mail_address`, do they have these names in other parts of the data?

```{r}
null_01 %>%
  filter(if_all(all_of(null_variable %>% pull(variable)),  ~is.na(.x))) %>% 
  select(employee_e_mail_address) %>% 
  distinct() %>% 
  inner_join(prep_base) %>% 
  select(params$unique_identifier_1,all_of(null_variable %>% pull(variable)))
```
